---
title: "STOR 565 Fall 2022 Homework 5"
author: "Aubrey Wang"
output:
  word_document: default
  html_document: default
  pdf_document: default
subtitle: \textbf{Due on 10/07/2022}
header-includes: \usepackage{amsgen,amsmath,amstext,amsbsy,amsopn,amssymb,mathabx,amsthm,bm,bbm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(ISLR)) { install.packages("ISLR", repos = "http://cran.us.r-project.org"); library("ISLR") }
if(!require(class)) { install.packages("class", repos = "http://cran.us.r-project.org"); library("class") }
if(!require(e1071)) { install.packages("e1071", repos = "http://cran.us.r-project.org"); library("e1071") }
if(!require(splines)) { install.packages("splines", repos = "http://cran.us.r-project.org"); library("splines") }
```
\theoremstyle{definition}
\newtheorem*{hint}{Hint}

\theoremstyle{remark}
\newtheorem*{rmk}{Remark}

*Remark.* Credits for **Theoretical Part** and **Computational Part** are in total *110 pt*. At most *100 pt* can be earned. For **Computational Part**, please complete your report in the **RMarkdown** file and submit your printed PDF homework created by it.

## Computational Part

>> **You are supposed to finish the Computational Part in a readable report manner for each dataset-based analysis problem. Other than what you are asked to do, you can decide any details on your own discretion.**

**Note.** Read Section 9.6 and 7.8 in Textbook before getting started.

1. (LDA \& QDA \& KNN \& SVM,*25 pt*) In Homework 4, we have performed serveral classification techniques to predict whether an area has high or low crime rate based on the `Boston` data set. We will continue with LDA, QDA, KNN and SVM in this exercise. Write a data analysis report addressing the following problems. We only use `indus`, `nox`, `age`, `dis`, `rad` and `tax` as the predictors. They are most associated with the binary response `crim01` as argued in Homework 4. 
 
    (a) Create a binary variable `crim01` as in Homework 4, and regenerate the training set and test set with ratio 1:1. 
```{r}
library(MASS)
library(tidyverse)
library(class)
data(Boston)
head(Boston)
library(caret)

median = median(Boston$crim)
my_data= mutate(Boston, crim01 = ifelse(crim >= median, 1,0))
my_data= my_data[c(3,5,7,8,9,10,15)]
head(my_data)
set.seed(100)
sample = sample(1:nrow(my_data),nrow(my_data)/2)
train  <- my_data[sample, ]
test <- my_data[-sample, ]
```
    (b) Perform LDA on the training data in order to predict `crim01` using the variables `indus`, `nox`, `age`, `dis`, `rad` and `tax`. What is the test error of the model obtained? Report the False Positive Rate and False Negative Rate on the test data.
```{r}
lda<- lda(crim01 ~., data=train)
lda
lda.class= predict(lda,test)$class
table(lda.class,test$crim01)
mean = mean(lda.class==test$crim01)


#test error
1-mean

#false-positive rate
4/(4+126)
#false-negative rate
39/(39+84)

#Thus, the test error is 0.1699605. The false positive rate is 0.03076923, and the false negative rate is 0.3170732.
```
    (c) Perform QDA on the training data in order to predict `crim01` using the variables that seemed most associated with `crim01` in (b). What is the test error of the model obtained? Report the False Positive Rate and False Negative Rate on the test data.
```{r}   
qda <- qda(crim01 ~., data=train)
qda
qda_pred <- predict(qda, test)$class
table(qda_pred, test$crim01)
mean1=mean(qda_pred == test$crim01)

#test error
1-mean1

#false positive rate
3/(3+127)

# false negative rate
27/(96+27)

#Thus, the test error for QDA is 0.1185771, and the false positive rate is 0.02307692, and the false negative rate is 0.2195122

```
    (d) Perform KNN on the training data, with several values of $K$, in order to predict `crim01`. Use only `indus`, `nox`, `age`, `dis`, `rad` and `tax`. Report the cross-validation errors associated with different values of $K$. Comment on your results.
```{r} 
trControl = trainControl(method  = "cv", number  = 5)
fit <- train(as.factor(crim01) ~ indus+nox+age+dis+rad+tax,
  method     = "knn",
  tuneGrid   = expand.grid(k = 1:10),
  trControl  = trControl,
  metric     = "Accuracy",
  data       = train)
fit
```
    (e) Fit a support vector classifier to the training data with various values of `cost`, in order to predict whether an area has high or low crime rate. Report the cross-validation errors associated with different values of this parameter. Comment on your results.
```{r}   
set.seed(1)
svm_train= data.frame(x=train[c(1,2,3,4,5,6)],y=as.factor(train$crim01))
tune = tune(svm,y~.,data = svm_train,kernel = "polynomial",ranges = list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
summary(tune)
tune$best.parameters

tune$best.performance
#Thus, the CV result shows the lowest training error is 0.1064615，which is when cost = 100.
```

    (f) Now repeat (e), this time using SVMs with radial and polynomial basis kernels, with different values of `gamma` and   `degree` and `cost`. Comment on your results.
```{r} 
tune_radial = tune(svm, y~., data = svm_train ,
kernel = "radial", ranges = list(
cost = c(0.01, 0.1, 1, 10, 100, 1000),
gamma = c(0.5, 1, 2, 3, 4) )
)
tune_polynomial <- tune(svm, y~.,data = svm_train,kernel = "polynomial",ranges = list(cost = c(0.01, 0.1, 1, 10, 100, 1000),
              degree = c(1:10)))

tune_radial
tune_polynomial
#Thus, as for svm with radical kernel, cost = 10 and gamma = 2 returns the smallest CV which is 0.05985; and as for svm with polynomical kernel. cost = 1000 and gamma = 3 returns the smallest error which is 0.08738462.
```
    (g) Make some plots to back up your assertions in (d), (e) and (f).
```{r} 
plot(tune)
plot(tune_radial)
plot(tune_polynomial)
```
    (h) Compare the test errors of LDA, QDA, the best tuned models for KNN, linear SVM, SVM with radial basis kernel, and SVM with polynomial basis kernel.
    
The test errors of LDA is 0.1699605, test error of QDA is 0.1185771, test error of KNN is 0.1064615, test error of SVM with radial basis kernel is 0.05984615 , test error of SVM with polynomial basis kernel is 0.08738462 . We can see that the SVM with radial basis kernel performs best.
***

**Hint.** In the lab, we used the `plot()` function for svm objects only in cases with $p = 2$. When $p > 2$, you can use the `plot()` function to create plots displaying pairs of variables at a time. Essentially, instead of typing
    
```{r eval=FALSE}
plot(svmfit, dat)
```

where `svmfit` contains your fitted model and `dat` is a data frame containing your data, you can type

```{r eval=FALSE}
plot(svmfit, dat, x1 ~ x4)
```

in order to plot just the first and fourth variables. However, you must replace `x1` and `x4` with the correct variable names. To find out more, type `?plot.svm`.

***

2. (Polynomial Regression and Regression Spline, *25 pt*) This question uses the variables `horsepower` (engine horsepower) and `displacement` (engine displacement) from the `Auto` data. We will treat `horsepower` as the predictor and `displacement` as the response. Write a data analysis report addressing the following problems.

    (a) Use the `poly()` function to fit a cubic polynomial regression to predict `displacement` using `horsepower`. Report the regression output, and plot the resulting data and polynomial fits.
```{r}
head(Auto)
poly=lm(displacement~poly(horsepower,3),data=Auto)
summary(poly)
ggplot(data=Auto, aes(horsepower,displacement)) +
        geom_point() +
        geom_smooth(method="lm", formula=y~poly(x,3))
```
    (b) Plot the polynomial fits for a range of different polynomial degrees (say, from 1 to 10), and report the associated residual sum of squares.
```{r}  
par(mfrow=c(3,3))
for (i in 1:10){
  poly.fit=lm(displacement~poly(horsepower,i),data=Auto)
  print(deviance(poly.fit))
  par(mfrow=c(5,2))
  print(ggplot(data=Auto, aes(horsepower,displacement)) +
        geom_point() +
        geom_smooth(method="lm", formula=y~poly(x,i)))
}
```  
    (c) Perform cross-validation or another approach to select the optimal degree for the polynomial, and explain your results.

```{r}  
library(boot)
for (i in 1:10){ 
  poly.train=glm(displacement~poly(horsepower,i),data=Auto) 
  cv.error=cv.glm(Auto,poly.train,K=10) 
  print(cv.error$delta)
}   

#Thus, the optimal degree is 4 and the model has returned the smallest cv error. Thus, we have determined our complexity to be degree=4.
```
    (d) Use the `bs()` function to fit a cubic B-Spline to predict `displacement` using `horsepower`. Report the output for the fit using four degrees of freedom without an intercept. This means there is one knot. How did you choose the knots? Plot the resulting fit.
```{r}  
bs = lm(displacement~bs(horsepower,df=4),data=Auto)
hplims <- range(Auto$horsepower)
hp.grid <- seq(from = hplims[1], to = hplims[2])
pred <- predict(bs, newdata = list(horsepower = hp.grid), se = T)
plot(Auto$horsepower, Auto$displacement, col = "gray")
lines(hp.grid, pred$fit, lwd = 2)
attr(bs(Auto$horsepower, df = 4), "knots")

# Thus, as for how I choose the knots, it will locate at the value that divide the data to two parts. The attr function shows that the knots is at horsepower=93.5.
```
    (e) Now fit a cubic B-Spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS. Describe the results obtained.
```{r}   
par(mfrow=c(2,5)) 
for (i in 4:13){
  bs.cubic=lm(displacement~bs(horsepower,df=i),data=Auto)
  hplims <- range(Auto$horsepower)
  hp.grid <- seq(from = hplims[1], to = hplims[2])
  pred <- predict(bs.cubic, newdata = list(horsepower = hp.grid), se = T)
plot(Auto$horsepower, Auto$displacement, col = "gray")
  lines(hp.grid, pred$fit, lwd = 2)
  print(deviance(bs.cubic))
}

#  Looking at the RSS, we can conclude that spline with 9 knots returns the smallest RSS.

```
    (f) Perform cross-validation or another approach in order to select the best degrees of freedom for a cubic B-Spline on this data. Describe your results. 
```{r} 
for (i in 4:13){
fit=model.frame(displacement ~ bs(horsepower, df=i), data=Auto) 
names(fit)=c("displacement", "bs.horsepower")
spline.fit = glm(displacement~bs.horsepower,data=fit)
cv.error=cv.glm(spline.fit,data=fit,K=10)
print(cv.error$delta[1])
}
#the cross validation result shows that when df=12 I have the lowest CV error of 1907.
```
    (g) Use the `ns()` function to fit a natural cubic spline to predict `displacement` using `horsepower`. Report the output for the fit using four degrees of freedom without an intercept. This means there are three internal knots. 
```{r} 
ns.cubic=lm(displacement~ns(horsepower,df=4),data=Auto)
hplims = range(Auto$horsepower)
hp.grid = seq(from = hplims[1], to = hplims[2])
pred = predict(ns.cubic, newdata = list(horsepower = hp.grid), se = T)
plot(Auto$horsepower, Auto$displacement, col = "gray")
lines(hp.grid, pred$fit, lwd = 2)
summary(ns.cubic)
attr(ns(Auto$horsepower, df = 4), "knots")

# As 3 knots divided the curve into four pieces. Three knots located at 75, 93.5, and 126 respectively, The natural spline returns a adjr2 of 0.82, which is good.
```
    (h) Now fit a natural cubic spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS. Describe the results obtained.
```{r} 
par(mfrow=c(2,5)) 
for (i in 4:13){
  ns.cubic=lm(displacement~ns(horsepower,df=i),data=Auto)
  hplims <- range(Auto$horsepower)
  hp.grid <- seq(from = hplims[1], to = hplims[2])
  pred <- predict(ns.cubic, newdata = list(horsepower = hp.grid), se = T)
  plot(Auto$horsepower, Auto$displacement, col = "gray")
  lines(hp.grid, pred$fit, lwd = 2)
  print(deviance(ns.cubic))
}

# Thus, looking through the output, we can find that df=13 gives me the lowest training RSS. (as 680400.9 is the smallest output)
```
    (i) Perform cross-validation or another approach in order to select the best degrees of freedom for a natural cubic spline on this data. Describe your results. 
```{r}    
for (i in 4:13){
fit=model.frame(displacement ~ ns(horsepower, df=i), data=Auto)
names(fit)=c("displacement", "ns.horsepower")
spline.fit=glm(displacement~ns.horsepower,data=fit) 
cv.error=cv.glm(spline.fit,data=fit,K=10) 
print(cv.error$delta[1])
}
# Thus, when df=10 I have the lowest CV error of 1860.973.
```
***